---
title: "DA2 Assignment 2 Appendix"
output: pdf_document
---

# Distributions, Correlations

```{r , include = FALSE , warning = FALSE}
# Data Loading and cleaning

library(tidyverse)
library(lspline)
library(fixest)
library(modelsummary)
library(kableExtra)
library(mfx)
library(ggpubr)
library(reshape2)



hotels_europe_price <- read_csv("https://osf.io/p6tyr/download")
hotels_europe_features <- read_csv("https://osf.io/utwjs/download")
data <- left_join(hotels_europe_price, hotels_europe_features, by = "hotel_id")
rm(hotels_europe_price,hotels_europe_features)


hotels <- data %>% 
  filter( !is.na( stars )) %>%
  filter( city_actual=="Barcelona" ) %>%
  filter( !is.na( distance )) %>%
  filter( !is.na( rating )) %>%
  filter( year == 2017, month == 12 , weekend == 0)  %>% 
  filter( accommodation_type  == "Hotel")%>% 
  filter( price  <= 4500)


hotels <- hotels %>% 
  mutate( logprice = log( price ))
          # ,
          # rating_cutoff = 0 +
          #   1 * as.numeric( hotels$rating_reviewcount >= 100 ) +
          #   1 * as.numeric( hotels$rating_reviewcount >= 1000 ) 


hotels$highly_rated <- ifelse(hotels$rating >= 4, 1, 0)


```



```{r , echo = FALSE , warning=FALSE , error=FALSE , message = FALSE}
hotels %>%
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value) ) +
  theme_bw()+
  facet_wrap(~key, scales = "free") +
  geom_histogram()




numeric_hotels <- keep( hotels , is.numeric )
cT <- round( cor( numeric_hotels , use = "complete.obs") , 2 )
cT[ upper.tri( cT ) ] <- NA
melted_cormat <- melt( cT , na.rm = TRUE)
melted_cormat


datasummary_skim( hotels )
```

Notably, we can see that prices have lognormal distribution. They are also easier to interpret in % differences.
TripAdvisor rating and custome rating have 0.84 correlation. TripAdvisor ratings hopefully largely come from customer ratings so there is likely a big overlap in the two which is why they explain each others variation.


# Lowess graphs

```{r , echo = FALSE , warning=FALSE }


chck_sp <- function( x_var , x_lab ){
  ggplot( hotels , aes(x = x_var, y = highly_rated)) +
    geom_point(color='red',size=2,alpha=0.6) +
    geom_smooth(method="loess" , formula = y ~ x )+
    scale_y_continuous(limits = c(0,1), breaks = seq(0,1,0.2))+
    labs(x = x_lab, y = "High-Low Rating") +
    theme_bw()
}


g1 <- chck_sp( hotels$nnights , "number of nights" )
g2 <- chck_sp( hotels$logprice , "logprice" ) 

association_figs <- ggarrange(g1, g2,
                       hjust = -0.6,
                       ncol = 2, nrow = 1)
association_figs

```

We can see no association for number of nights spent by customer. For logprices, I decided to add knots at ln price = 5 and ln price = 5.6.


# Histogram of 2 nonlinear models' predictions



```{r , echo = FALSE , warning=FALSE , out.width="50%"}
probit <- glm(highly_rated ~ stars + lspline( distance , c( 1 , 2 ) ) 
              + lspline( logprice , c( 5 , 5.6 ) )
              , data = hotels , family=binomial(link="probit"))

logit <- glm( highly_rated ~ stars + lspline( distance , c( 1 , 2 ) ) 
             + lspline( logprice , c( 5 , 5.6 ) )
             , data = hotels , family='binomial')

lpm <- feols( highly_rated ~ stars + lspline( distance , c( 1 , 2 ) ) 
               + lspline( logprice , c( 5 , 5.6 ) )
               , data = hotels , vcov = "hetero")


hotels$pred_probit<- predict.glm(probit, type="response") 
hotels$pred_logit <- predict.glm(logit, type="response")
pred_lpm <- stats::predict(lpm)


ggplot( hotels , aes(x = pred_lpm)) +
  geom_histogram( binwidth = 0.1, fill='navyblue', color = 'white' ) +
  labs(y = 'Count',x = "Logit predictions") +
  theme_bw()

ggplot( hotels , aes(x = pred_probit)) +
  geom_histogram( binwidth = 0.1, fill='navyblue', color = 'white' ) +
  labs(y = 'Count',x = "Logit predictions") +
  theme_bw()

ggplot( hotels , aes(x = pred_logit)) +
  geom_histogram( binwidth = 0.1, fill='navyblue', color = 'white' ) +
  labs(y = 'Count',x = "Probit predictions") +
  theme_bw()


```

As intended nonlinear models don't have predictions above 1.