---
title: "Appendix"
output: pdf_document
---

```{r , include = FALSE , warning = FALSE}
# Data Loading and cleaning

rm(list=ls())
library(tidyverse)
library(modelsummary)
library(kableExtra)
library(ggpubr)
library(fixest)
library(xtable)


cps_earnings <- read_csv("https://osf.io/4ay9x/download")


cps_earnings_filtered <- cps_earnings %>% 
  filter( grade92 >= 44 & grade92 <= 46 ) %>%
  filter( age >= 24 & grade92 <= 65 ) %>% 
  filter( lfsr94 == "Employed-At Work" ) %>%
  filter( uhours >= 20 ) %>% 
  filter( earnwke > 5 )


cps_earnings_filtered <- cps_earnings_filtered %>% 
                         mutate( female =  as.numeric( sex == 2 )) %>%
                         mutate( wages_per_hour = earnwke / uhours) %>%
                         mutate( log_wages_per_hour = log( wages_per_hour )) 

cps_earnings_filtered_teach <- cps_earnings_filtered %>%
          mutate( sample = ifelse( occ2012 == 2200 , 1 , 0 ) ) 

earnings_teachers <- cps_earnings_filtered_teach %>% filter( sample == 1 )



```

## Loess graphs


```{r , echo = FALSE , warning = FALSE, out.width = "75%"}
# Lowess graphs
p1 <- ggplot( data = earnings_teachers , aes( x = female , y = log_wages_per_hour )) +
  geom_point( color = 'red', size = 2, shape = 16, ) + 
  scale_x_continuous(expand=c(0.01, 0.01), limits = c(0, 1),   breaks=seq(0, 1,   by=1)) + 
  scale_y_continuous(expand=c(0.01, 0.01),limits = c(0, 5), breaks=seq(0, 5, by=1)) +
  geom_smooth( method = 'loess', formula = y ~ x) +
  labs(x = "Gender",y = "Log of wages (USD)")


p2 <- ggplot( data = earnings_teachers , aes( x = grade92, y = log_wages_per_hour )) +
  geom_point( color = 'red', size = 2, shape = 16, ) + 
  scale_x_continuous(expand=c(0.01, 0.01), limits = c(44, 46),   breaks=seq(44, 46,   by=1)) + 
  scale_y_continuous(expand=c(0.01, 0.01),limits = c(0, 5), breaks=seq(0, 5, by=1)) +
  geom_smooth( method = 'loess', formula = y ~ x) +
  labs(x = "Level of Education",y = "Log of wages (USD)")



linear_assoc_figs <- ggarrange(p1, p2,
                              hjust = -0.6,
                              ncol = 1, nrow = 2)
linear_assoc_figs


```

Based on the non-parametric regression lines, it seems that the association between gender and log wages is linear. For education and log wages, the overall picture is also linear, however, there is a change in the trend between Professional degrees and PhD, possibly because we have a smaller number of Professional degress in our sample.

A linear spline model with knot at Professional degress might produce a better fit modell, which could be useful for predicting wages. However, for the overall association of the data this is likely not significant as we show in multiple regression model with education as Categorical variable.